---
title: "Coursera Practical Machine Learning Assignment (Yee Ling, Foo)"
author: "Foo Yee Ling"
date: "Saturday, June 21, 2014"
output: html_document
---

This assignment requires the use of sensors data from 6 fitness enthusiasts to build a model to classify the correct and incorrect use of the fitness equipment.

The codes run on R using the Caret and Kernlab package.

```{r}
library(caret)
library(kernlab)
set.seed(327)
```

Import the full data from its source (http://groupware.les.inf.puc-rio.br/har) and use a subset of the training data that gives the sensors readings and all the outcome for the 5 different classes. Select 8000 rows out of the 39242 observations to build and validate the model.

```{r}
pmldataA = read.csv(file="pml-training-all1.csv")
pmldata = pmldataA[sample(nrow(pmldataA),size=8000,replace=FALSE),]
```


To start the modeling process, splice the data with 70% of it to be used for training and the remaining for prediction and testing.


```{r}
inTrain = createDataPartition(y=pmldata$classe, p=.7, list=FALSE)

pmlTr = pmldata[inTrain,] ## training data
pmlTt = pmldata[-inTrain,] ## testing data

dim(pmlTr)
dim(pmlTt)
```

In view of the large number of predictors in the data, a Principal Component Analysis (PCA) can be done as a pre-processing step to reduce the number of predictors.  

```{r}
## use pca
preProc = preProcess(pmlTr[,-53], method="pca" ) 
preProc
```

In this dataset, 26 components are kept to capture 95 percent of the variance.

A Support Vector Machine (SVM) classification model is then applied on the reduced training dataset using a 10-fold cross-validation.

```{r}
pmlTrPCA = predict(preProc, pmlTr[,-53]) ## PCA treated training data
pmlModelPCA = train(pmlTr$classe ~., method = "svmRadial", tuneLength = 10, preProc = c("center","scale"), trControl = trainControl(method = "cv"), data = pmlTrPCA)
plot(pmlModelPCA, metric="Accuracy")
```

With the use of cross-validation, the accuracy of the the model is high at 95% and above.

Next, to find the out-of-sample errors, the test data is pre-processed with the  PCA model and the outcome is predicted using the SVM model.

```{r}
pmlTtPCA = predict(preProc, pmlTt[,-53]) ## PCA treated test data
predClasse = predict(pmlModelPCA, pmlTtPCA) ## apply the model on the treated data
confusionMatrix(predClasse,pmlTt$classe)
```

Using the confusion matrix to validate the prediction results, and it can be seen that the svm model works very well, yielding approximately 96% accuracy. The sensitivity (recall) and specificity (precision) for the 5 classes are high at more than 93%. The 10-fold cross validation has been very balanced with more than 96% accuracy, thus demonstrating a robust and reliable model which can be confidently applied onto the 20 test cases. 

```{r}
pmldataT = read.csv("pml-testing1.csv")
pmlTtPCA20 = predict(preProc, pmldataT[,-53])
predict(pmlModelPCA, pmlTtPCA20)
```

